---
title: "PhyloSeq"
author: "Jair"
date: "2024-04-18"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(phyloseq)
library(dplyr)
library(ggplot2 )
library(BiMiCo)
library(devtools)
library(MicEco)
library(vegan)
```

```{r}
load("RData/seqtab.nochim.RData")
load("RData/taxa.RData")
```

```{r}
metadata<-read.csv("sample-metadata-phylo.csv", header=TRUE)
head(metadata)
```
```{r}
metadata <- read.csv("sample-metadata-phylo.csv", header=TRUE)
row.names(metadata) <- metadata$sample.id  # Set sample IDs as row names
# Assuming your OTU table and taxonomic data are already correctly loaded
otu_table_object <- otu_table(seqtab.nochim, taxa_are_rows=FALSE)
tax_table_object <- tax_table(taxa)  # Ensure this is loaded as you haven't shown this part

# Create the phyloseq object
physeq <- phyloseq(otu_table_object, sample_data(metadata), tax_table_object)
print(physeq)

```

```{r}
# Check metadata sample names
sample_names(sample_data(metadata))
```
#transform sample counts
```{r}
#convert from raw to abundance so its easier to comapre
physeq <- transform_sample_counts(physeq, function(abund) 1*(abund>0))
physeq
```

```{r}
dna <- Biostrings::DNAStringSet(taxa_names(physeq))
names(dna) <- taxa_names(physeq)
physeq <- merge_phyloseq(physeq, dna)
taxa_names(physeq) <- paste0("ASV", seq(ntaxa(physeq)))
physeq
```
#Exclude non-bacterial sequences
```{r}
physeq <- physeq %>% subset_taxa( Family!= "Mitochondria" | is.na(Family) & Order!="Chloroplast" | is.na(Order)) 
```

```{r}
physeq <-rm_nonbac(physeq)
physeq
```

```{r} 
save(physeq, file= "RData/physeq.RData")
```

```{r}
load("RData/physeq.RData")
```

#April 23

```{r}
plot_bar(physeq, fill = "Phylum") + geom_bar(aes(color=Phylum, fill=Phylum),
stat="identity", position="fill")
```

```{r}
physeq_relabund <- transform_sample_counts(physeq, function(x) x / sum(x))
```

#barplot 
```{r}
plot_bar(physeq_relabund, fill = "Phylum") + geom_bar(aes(color=Phylum, fill=Phylum), stat="identity", position="fill") + facet_wrap(~body.site, scales="free")
```

#plot alpha diversity based on body.site
```{r}
plot_richness(physeq, x="body.site", color= "subject", measures=c("Observed", "Simpson", "Shannon"))
```

#plot alpha diversity based on subject
```{r}
plot_richness(physeq, x="subject", color= "subject",
measures=c("Observed", "Simpson", "Shannon"))
```

```{r}
alpha <- estimate_richness(physeq, measures=c("Observed", "Simpson", "Shannon"))
```

```{r}
#shapiro-wilk
observed <- shapiro.test(alpha$Observed)
simpson <-shapiro.test(alpha$Simpson)
shannon <-shapiro.test(alpha$Shannon)

print(observed)
print(simpson)
print(shannon)
```

#create data frames for stats analysis
```{r}
samples <-sample_data(physeq)

if (class(samples) == "sample_data") {
  samples <-data.frame(sample_data(samples))
}

alpha$sample <-rownames(alpha)
alpha <-merge(alpha, samples, by = "sample")
```

```{r}
test_observed <- wilcox.test(Observed ~ subject, data = alpha)
test_simpson <- wilcox.test(Simpson ~ subject, data = alpha)
test_shannon <- wilcox.test(Shannon ~ subject, data = alpha)
```

```{r}
kruskal.test(Simpson ~ body.site, data=alpha)
```
```{r}
pairwise.wilcox.test(alpha$Simpson, alpha$body.site, p.adjust.method = "holm")
```
```{r}
kruskal.test(Observed ~ body.site, data=alpha)
```
```{r}
pairwise.wilcox.test(alpha$Observed, alpha$body.site, p.adjust.method = "holm")
```
```{r}
shannonanova <- aov(Shannon ~ body.site, data=alpha)
summary(shannonanova)
```
```{r}
TukeyHSD(shannonanova)
```
#Significant differences between tongue-gut, tongue and both palms

#4/25
#Remove tax with relative abundance <0.005%
```{r}
minTotRelAbun = .00005
x = taxa_sums(physeq)
keepTaxa = (x/sum(x)) > minTotRelAbun
physeqprune = prune_taxa(keepTaxa, physeq)
physeqprune
```

#Number of shared ASVs body.site )found in 25% or more)
```{r}
bodysite=ps_venn(
  physeqprune,
  "body.site",
  fraction = .25,
  weight = FALSE,
  relative = TRUE,
  plot = TRUE
)
```

#Bray Curtis Calculation
```{r message = FALSE}
set.seed(777)
dist = phyloseq::distance(physeqprune, method="bray", weighted=TRUE)
ordination = ordinate(physeqprune, method = "NMDS", distance=dist)
```

#Bray Curtis Subject Plot
```{r}
braysubject=plot_ordination(physeqprune, ordination, color="subject") + 
    theme_classic() +
  theme(strip.background = element_blank()) + stat_ellipse(aes(group=subject))
braysubject
```

#Bray Curtis Antibiotic Usage Plot
```{r}
brayabusage=plot_ordination(physeqprune, ordination, 
color="reported.antibiotic.usage") + 
  theme_classic() + 
  theme(strip.background = element_blank()) + stat_ellipse(aes(group=reported.antibiotic.usage))
brayabusage
```

```{r}
adonis2(dist ~ sample_data(physeqprune)$subject)
```

```{r}
adonis2(dist ~ sample_data(physeqprune)$body.site)
ps.disper<-betadisper(dist, sample_data(physeqprune)$body.site)
permutest(ps.disper, pair = TRUE)
```

```{r}
ggsave(
  filename = "Fig1.tiff",
  plot = brayabusage,
  width = 200,  # assuming you meant to use mm and this size is intentional
  height = 200,
  units = "mm",
  dpi = 300
)
```